---
title: "Week 7: Analysing Quantitative Data"
format: html
editor: source
execute: 
  eval: FALSE
  echo: TRUE
  warning: FALSE
  message: FALSE
---

# Lecture 1: Choosing a Statistical Test {.unnumbered}

```{r}
#| eval: TRUE


library(tidyverse)

cows <- readxl::read_excel("cows.xlsx") |> 
  mutate(parlour = as.factor(parlour),
         `housing type` = as.factor(`housing type`))


cows |> 
  ggplot(aes(x = `Welfare Score`, y= `Average Daily Yield`, colour = `housing type`)) +
  geom_point() +
  theme_classic() +
  labs(x  = "Welfare Score", y = "Average Daily Yield",
       title = "Milk Yield versus Welfare Score",
       subtitle = "For Robotic vs Manual Parlours") +
  facet_wrap(facets = ~parlour, ncol = 1) +
  theme(legend.position = "bottom")

```

## Numerical Response {.unnumbered}

```{r}
#| eval: true
cows |> 
  summarise(mean_yield = mean(`Average Daily Yield`),
            sd_yield = sd(`Average Daily Yield`))


t.test(cows$`Average Daily Yield`, mu = 28)

```


## Numerical Response Categorical Explanatory{.unnumbered}

```{r}
#| eval: true



diet <- tibble (before = c(5.04, 4.63, 4.04, 5.10, 5.43, 4.83, 3.45, 3.49, 5.02, 4.81),
                after = c( 4.78, 2.49, 4.46, 2.03, 5.13, 7.23, 3.50, 1.89, 3.30, 3.91))


diet |> 
  ggplot(aes(y = before)) +
  geom_boxplot() +
  geom_boxplot(aes(y = after, x =1)) +
  theme_classic() +
  scale_x_continuous(labels =c("Before Diet", "After Diet"), breaks = c(0,1)) +
  labs(x = "Diet", y = "Weight (kg)")



diet |> 
  summarise(before_mean = mean(before),
            after_mean = mean(after),
            before_sd = sd(before),
            after_sd = sd(after))


t.test(diet$before, diet$after, paired = TRUE, alternative = "two.sided")


```



## Numerical Response Numerical Explanatory {.unnumbered}



## Categorical Response {.unnumbered}




## Categorical Response Categorical Explanatory{.unnumbered}




## Categorical Response Numerical Explanatory {.unnumbered}


## Parametric vs Non Parametric {.unnumbered}

## Power Calculations {.unnumbered}

## Assumptions {.unnumbered}